Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-05-10 research_proposal/research_proposal_draft.qmd
super small fix
final touch

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
execute:
  cache: true
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
---

@@

However, the hierarchical time series data often exhibit a prominent principal components structure, which is not taken into account by the shrinkage estimator. Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions. We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$. The twenty largest eigenvalues of the covariance matrix, which we scale by logarithm, are plotted in @fig-eigen. We can see that the first 4 eigenvalues are significantly larger than the others, indicating a prominent principal components structure.

![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/tourism_eigen.png){#fig-eigen width="70%"}

@@

\For{$i = v:T-1$ do}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
    \EndFor
\EndFor

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: Vincent Su
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
execute:
  cache: true
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
---

@@

However, the hierarchical time series data often exhibit a prominent principal components structure, which is not taken into account by the shrinkage estimator. Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions. We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$. The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen. We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.

![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}

@@

\For{$i = v:T-1$}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
    \EndFor
\EndFor
